An Audio Interface for an Active Perception Navigation Aid for Blind People

### 2019-05-13 article.tex
Fixed formatting for new template. Starting great purge

\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}

@@

% Metadata Information
\acmJournal{TAP}
%\acmVolume{9}
%\acmNumber{4}
%\acmArticle{39}
%\acmYear{2010}
%\acmMonth{3}
%\acmArticleSeq{11}

@@

To this end, we are developing a mobile device-based navigation system using a Google Project Tango device\footnote{https://developers.google.com/tango}, pictured in Fig.~\ref{fig:tango-headphone}, that caters to the perception limitations of people with limited vision by using a non-visual interface. This system will ultimately allow users with visual impairments to reach a target location in the final leg of their journey, i.e.\ the so-called `last 10 yards' problem~\cite{google2016blind,bellotto2013}. The ideas here can also be extended and applied to existing macronavigation tools, such as Google Maps. A Tango-enabled device comes pre-equipped with powerful image-processing, localisation and depth-perception capabilities and is built on top of a standard Android platform, providing access to the entire set of input/output options that this operating system has to offer. The final system will use multiple feedback modes (vibration, audio and voice) to guide a user toward a target destination while providing information on any oncoming obstacles.

\begin{figure}
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=0.85\textwidth]{figures/tango_headphone.png}
    \caption{The Tango device and bone-conducting headset used by the participants. }
    \label{fig:tango-headphone}
  \end{subfigure}\hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=0.7\textwidth]{figures/camera_coordinate.pdf}
    \caption{A diagram showing the Tango's frame of reference and the target's pan ($\phi$) and tilt ($\theta$) angles.}
    \label{fig:cam-coords}
  \end{subfigure}
  \caption{The hardware used during the experiment, as well as a diagram showing the Tango device's coordinate system. }
\end{figure}

In this paper we focus on the audio feedback mode and discuss how it is used in our system. We present experiments we performed to determine its effectiveness at directing a user to complete a pointing task and how its parameter values affect a user's target-search performance. For the pointing task, we presented blindfolded and partially sighted participants with a set of virtual targets, one after another, and asked them to simply point a camera to where they thought the targets were. The targets' pan and tilt angles w.r.t.\ the Tango's vertical plane, shown in Fig.~\ref{fig:cam-coords}, were given to the participants through a spatial tone with varying pitch played via a set of bone-conducting headphones. We use the latter to not interfere with the user's normal hearing function, which people with visual impairments typically rely on. Because these headphones bypass the external structure of the ear responsible for localising a sound source's elevation, it is necessary to convey the target's tilt angle using another method. Therefore, we opted for a tone with varying pitch to convey the angle adjustment required to point the camera in a particular direction.

@@

%####################################################
REVIEW SECTION - MAKE SURE TO MENTION DIFFERENCE BETWEEN BLIND AND SIGHTED , sonification of 3d scenes in an electronic travel aid for the blind.
Spatial audio has also been considered to convey the direction of a target. Experiments have previously determined that people are able to find the location of a stationary sound source with an error of $\pm$\SI{35}{\degree}, in both the pan and tilt dimensions~\cite{zwiers2001spatial}, for both early-blind and normally-sighted people. However, in~\cite{lewald2013exceptional, lessard1998early} it was found that the blind have a clear advantage in localisation accuracy over sighted people when presented with more difficult tasks, such as targets in motion and narrow-band stimuli, with~\cite{lewald2013exceptional} reporting an average absolute localisation error of around \SI{10}{\degree} in the pan and tilt dimensions. Other authors found that the minimum difference in the spatial sound's angle for the user to be able to perceive movement is approximately \SI{1.7}{\degree}. During the experiments, an audio speaker was physically manoeuvred to provide the participant with a spatialised sound tone~\cite{ashmead1998spatial}. Researchers have also tried using simulated spatial audio to inform the user from which direction a sound is originating~\cite{holland2002audiogps, kammoun2012navigation, rebillat2009smart, menelas2010audio, wilson2007swan, zotkin2004rendering}. In these works, a sound was played through a set of headphones and the source spatialised with a head-related transfer function (HRTF), tricking the listener into thinking the sound source was located at some arbitrary 3D location. Various audition techniques and methods were then used to guide the user along a path. Other authors have also provided a framework for evaluating the quality of this 3D sound in terms of its psychoacoustic properties~\cite{guastavino2004perceptual, nicol2014roadmap}. There are also experimental results about how well users can find targets presented with spatial sound in the tilt and panning dimensions with normal headphones~\cite{katz2011spatial, zwiers2001spatial}, as well as work that has shown similar levels of localisation performance between normal and bone-conducting headphones in the pan dimension~\cite{macdonald2006spatial}. However, to our knowledge no extensive work or experiments have been done to determine how well users respond to tilt adjustment instructions using a tone with \emph{varying pitch}, in particular when using bone-conducting headphones. 
%########################################################

@@

We use a set of bone-conducting headphones (see Fig.~\ref{fig:tango-headphone}) that are placed externally on the user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function. In the future, our system will use multiple feedback modes to provide the user with navigation and obstacle avoidance instructions. However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.

A diagram of the experimental system pipeline is shown in Fig.~\ref{fig:pipeline}. Here, the arrows indicate the direction of the flow of information. When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the Tango's current position and orientation. The audio generator then produces a tone based on the difference between the Tango and the target's positions (in the Tango's global coordinate system) and sends it to the audio output channel, which plays it back to the user. The WiFi recording module is constantly monitoring the different parameter values of the Tango and target's positions, as well as the system's output, and records it to a remotely stored datafile.

\begin{figure}
  \centering
  \includegraphics[clip=true, trim=50 120 80 50, width=0.55\textwidth]{figures/pipeline.pdf}
  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }
\label{fig:pipeline}
\end{figure}

@@

For the series of experiments performed in this work, only the audio feedback mode was used to interface with the user. The audio component is responsible for conveying the 2D position of a target on the vertical plane w.r.t.\ the Tango's coordinate system, showed in Fig.~\ref{fig:cam-coords}, in terms of pan and tilt angles. Distance to the target can be rendered by adjusting the audio signal's amplitude as a function of the distance. However, in this work, this was not implemented. The audio signal is a sinusoidal sound wave that is continuously played (refreshed at $\sim$\SI{10}{\hertz}) to the user through bone-conducting headphones. We select a sinusoidal wave because it is relatively simple to manipulate and analyse. A richer and more natural tone will be implemented at a later stage of the project.

@@

The pan angle describes the angle which the user needs to rotate the camera vector around the vertical $z$-axis in Fig.~\ref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user. We use an HRTF to add a spatial element to the audio tone that the system plays to the user, making the tone sound like it is coming from the direction of the target. Humans determine the pan direction by subconsciously analysing the time difference between a sound reaching both of your ears. For example, if you hear a sound in your left ear first, you interpret the sound as coming from your left. The greater the so-called inter-aural time difference~\cite{wightman1992dominant}, the greater the perceived angular distance to the sound source.

@@

The system adjusts the tone's pitch to convey the target's tilt angle w.r.t.\ the camera's current pointing vector, $C$, as shown in Fig.~\ref{fig:cam-coords}. Here, a high pitch means the target is above the camera vector (i.e.\ the user should look up) and a low pitch means the target is below the camera vector (the user should look down). This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial, blauert1997spatial}. We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.

@@

The frequency is selected using Equation~\ref{eq:pitch-eq}, where $\theta$ is the tilt angle and $m$ and $c$ are parameters calculated using a linear equation (seen in Equation~\ref{eq:linear}) and the cut-off frequency values. Here, $y_m$ is a cut-off frequency, $x_m$ its corresponding angle and $m$ the gradient for the setting, calculated as $m = \frac{\Delta y}{\Delta x}$. The pitch gain functions are shown in Fig.~\ref{fig:pitch-gain}.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.55\textwidth]{figures/pitch_gain_functions.png}
  \caption{The pitch gain functions for the 3 different gain settings. }
  \label{fig:pitch-gain}
\end{figure}

@@

For the experiments we used two groups of participants, with one including 42 sighted and blindfolded participants and the other 11 legally blind volunteers with varying degrees of vision impairment. We refer to these 2 groups as G1 and G2.
%(see Fig.~\ref{fig:participants} for pictures of two participants during the experiments)
The participants of each group performed a series of experiments using our system with a pair of bone-conducting headphones.
The participants were recruited on a volunteer-basis and consisted of a diverse group of undergraduate students, professionals and retirees with ages ranging between 17 and 95 years (18 male, 35 female).
Some of the participants were using hearing aids.
However, this was not an issue, since we used bone-conducting headsets which bypass the outer ear structure and the hearing aids.

The participants took part in three experiments.
The first two were performed to determine each participant's hearing characteristics and provide some context to the results generated during the target-search experiment.
These characterisation experiments also indicate any pre-existing biases the participant-base might have in their hearing behaviour for our third, target search experiment.

@@

The distance between a participant and the target is not considered here and the 3D targets are therefore generated at a constant distance of \SI{2}{\metre} from the participant.
Throughout the experiments, various parameters of the target and the participant are recorded and streamed in real-time to a laptop computer via a WiFi connection.
Prior to the target search experiment, participants from G1 were given a few minutes without a blindfold to familiarise themselves with the system and the audio tone's behaviour as the target's position changes.
This was done verbally with the participants from G2 to inform them of the pitch they were aiming for that indicated that they were on-target.

@@

In this experiment, we evaluated a participant's ability to determine the lateral direction a sound is coming from.
To do this, we played a continuous \SI{512}{\hertz} sinusoidal tone to the participant through the headphones and applied an HRTF to spatialise and place its source to the participant's left or right.
The participant then had to select the direction the sound came from.
The longer the experiment lasts and the more correct guesses the participant makes, the closer the source moves to the centre-front of the participant, making it increasingly difficult to distinguish between the left and right direction of the sound.

For this progressive increase in difficulty, a 2-up, 1-down step procedure is used~\cite{wetherill1965sequential,levitt1971transformed}, meaning that for every 2 correct answers, the distance to the centre-front halves, making the lateral direction harder to determine.
Conversely, the task becomes easier for each incorrect answer by doubling the sound source's distance from the centre-front.
We also set  different step sequences: one starting at a large angular distance (\SI{45}{\degree}) from the participant's centre-front and the other at the minimum angular distance (approximately \SI{1}{\degree}), giving an `easy' and a `hard' step respectively.
The terminating condition for the experiment is when the hard and easy step sequences converge to within 2 steps of one another (i.e.\ the difference between the hard and easy sequences are 4 times larger or smaller than one another) for 3 consecutive guesses.
This gives an angular distance band which indicates each participant's ability to localise the sound source's lateral position.
Each participant performed this experiment three times.

@@

Next we determined each participant's ability to differentiate an audio tone's pitch, i.e.\ how well they can tell if a tone is high or low pitched.
We played 2 tones after another to each participant, with the second tone being higher or lower-pitched than the first tone.
The participants then had to indicate whether they thought the second tone as higher or lower-pitched than the first.

The first tone is randomly generated, while the second tone is generated by adding or subtracting some value from the first.
This tone difference depends on how well the participant can tell the tones apart.
Like the spatial awareness experiment, a 2-up, 1-down step process is used: for every 2 consecutive correct answers, the pitch difference between the tones is halved, and the difference is doubled for every incorrect answer.
Two step sequences are again used here, one starting with a large pitch difference ($f_h=2^9=$ \SI{512}{\hertz}) between the tones and the other with a small difference ($f_l=2^1=$ \SI{2}{\hertz}). 
The termination condition is when the two step sequences are within one octave of each other (i.e.\ $\log_2\frac{f_h}{f_l}=2^1$) for 3 consecutive answers.
Each participant performed this experiment twice.

@@

The final experiment is the main one and will answer the question we are most interested in: how well does a spatial tone with varying pitch direct a person to look in a specific direction, and how do the parameters of this tone affect a person's performance in this task?

For this experiment, the participants given a Tango device running an app purpose-written for this experiment.
When started, a set of virtual targets were presented one at a time to each participant on the Tango device.
Then, depending on the direction the participant was currently pointing the camera relative to the target's position, the Tango generated and played an audio tone via bone-conducting headphones to indicate the pan and tilt angle adjustment required to point the camera to the target.
%These instructions are spatialised tones with varying pitch: an HRTF indicates the pan direction (left or right) and the pitch indicates whether the participant should be looking up (high pitch) or down (low pitch) to find the target.

Once a participant pointed the Tango's camera toward the target, the system generated a neutral tone with a pitch of \SI{512}{\hertz}, which we used as the `on-target' pitch for all of our experiments.
A participant then had to manually confirm the direction they believed the target was after hearing the tone's pitch change to the \SI{512}{\hertz}, on-target tone, and tap the screen to confirm the location.
At this point a new search target was presented and the process was repeated.
To minimise any speed/accuracy trade-off bias in the results, we asked the participants to search for the targets without placing particular emphasis on accuracy or time.

A total of 28 targets were presented to each participant per experiment run.
The positions of these targets were randomly generated and equally spread across the 4 quadrants on the vertical plane to prevent a clustering of targets at the same location.
After every experiment run, the parameters controlling the tone's behaviour were adjusted.
In this case, the rate of change of the tone's pitch was adjusted to make the pitch increase at a lower or higher rate as a function of the tilt angle between the target and the participant's current pointing direction.
This was done to determine whether, for example, a more rapid increase in pitch would help the participant find the target faster or more accurately. 
The order in which the settings were set for each participant was randomised in order to minimise any learning effect the participant population might display between experiments.

@@

We use 2 different metrics to compare the 3 different pitch gradient settings, namely target acquisition accuracy and search time.
The accuracy is given as the difference between the location the Tango's camera was pointing at the time a participant selected a target's location and the target's actual position.
We separate the results of the tilt and pan dimensions in order to see how the different pitch gradients affect a participant's pointing accuracy.

We also compare the performance of the 3 pitch gradient settings in terms of the time it took each participant to find a target.
However, since each participant was presented with a different, randomly generated set of targets, a direct time comparison is not possible.
Instead, we use Fitts's Law~\cite{fitts1954information}, modified by MacKenzie for uncertain target sizes and noisy data~\cite{mackenzie1992fitts}, which states that there is a relation between the time it takes to find a target and its index of difficulty (the ratio between the distance to the target and its width).
It also provides us with an additional metric called the `index of performance' that we can use to compare the results between the 3 configurations. 
Here we briefly summarise the equations and quantities involved in this metric. 
Fitts's Law is given by

@@

\noindent
where $t$ is the time it takes to find a target, $a$ and $b$ are constants determined through regression and $ID$ is a target's index of difficulty, given as logarithmic function of the ratio between the distance to the target and the target's width.
In this case, the targets have no width since they are points in space, and we use MacKenzie's modified form for $ID$, given by

@@

\noindent
where $\sigma$ is the standard deviation of the error data, taken as the angle between a participant's target selection and the target's actual position. 
Fitts's index of performance, $IP$, can then be calculated with

@@

Fig.~\ref{fig:location-guesses} shows the results for the spatial awareness experiment in the pan direction where each participant determined the location of the sound they were played.

\begin{figure}
  \centering
  \includegraphics[clip, trim=100 20 150 100, width=0.5\textwidth]{figures/location_guesses_sighted.png}
  \caption{Histograms of the spatial awareness experiment in the pan dimension showing the proportion of correct/incorrect guesses per bin interval. }
  \label{fig:location-guesses}
\end{figure}

MEME INSERT BLIND FIGURE

With a relatively low number of erroneous guesses, we see that the participants were accurate in guessing the location of the sound source.
This is further supported by the large number of samples at the minimum difference of \textasciitilde\SI{3}{\cm}, indicating that the participants reached this level more frequently and consistently. We can therefore conclude that the participants had little difficulty determining the pan angle of a sound source. These results are in line with what we expected and they are supported by literature which indicates that humans are very adept at localising a sound source, and this ability was apparent for the HRTF-generated pan locations.

The results recorded during the pitch discrimination experiment are shown in Fig.~\ref{fig:tone-guesses}, where a bar plot is used to show the number of correct guesses for each tone difference level. The latter is given in terms of the relative semitone difference between the tones which are obtained using the following relation:

@@

From Fig.~\ref{fig:tone-guesses-sighted}, we can see that the guesses are normally spread around a 0 semitone difference for the blindfolded participants in G1, with roughly an equal number of samples with positive and negative difference values.
As expected, a larger proportion of incorrect guesses occur around the smaller differences with the majority of samples for correct and incorrect guesses being recorded in the $[-0.25, 0.25]$ semitone difference level.
However, the results from the participants from G2 in Fig.~\ref{fig:tone-guesses-blind} are less evenly spread and more concentrated around the 0 semitone difference level, with a skewing of wrong guesses values towards the positive side, possibly indicating a perception bias amongst our pool of participants with vision impairments. However, one possible reason for this bias may simply be a lack of data, and we may see that more samples from larger group of participants with vision impairments in G2 would provide a more normal spread, similar to the plot in Fig.~\ref{fig:tone-guesses-sighted}.

>>>>

\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{todonotes}
\usepackage[capitalise]{cleveref}

@@

% Metadata Information
%\acmConference{SAP}
%\acmVolume{9}
%\acmNumber{4}
%\acmArticle{39}
%\acmYear{2010}
%\acmMonth{3}
%\acmArticleSeq{11}

@@

To this end, we are developing a mobile device-based navigation system using a Google Project Tango device\footnote{https://developers.google.com/tango}, pictured in \cref{fig:tango-headphone}, that caters to the perception limitations of people with limited vision by using a non-visual interface.
This system will ultimately allow users with visual impairments to reach a target location in the final leg of their journey, i.e.\ the so-called `last 10 yards' problem~\cite{google2016blind,bellotto2013}. 
The ideas here can also be extended and applied to existing macronavigation tools, such as Google Maps.
A Tango-enabled device comes pre-equipped with powerful image-processing, localisation and depth-perception capabilities and is built on top of a standard Android platform, providing access to the entire set of input/output options that this operating system has to offer.
The final system will use multiple feedback modes (vibration, audio and voice) to guide a user toward a target destination while providing information on any oncoming obstacles.

%\begin{figure}
  %\centering
  %\begin{subfigure}[t]{0.45\textwidth}
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figures/tango_headphone.png}
  \caption{The Tango device and bone-conducting headset used by the participants. }
  \label{fig:tango-headphone}
\end{figure}
  %\end{subfigure}\hfill
  %\begin{subfigure}[t]{0.45\textwidth}
\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figures/camera_coordinate.pdf}
  \caption{A diagram showing the Tango's frame of reference and the target's pan ($\phi$) and tilt ($\theta$) angles.}
  \label{fig:cam-coords}
\end{figure}
  %\end{subfigure}
  %\caption{The hardware used during the experiment, as well as a diagram showing the Tango device's coordinate system. }
%\end{figure}

In this paper we focus on the audio feedback mode and discuss how it is used in our system.
We present experiments we performed to determine its effectiveness at directing a user to complete a pointing task and how its parameter values affect a user's target-search performance.
For the pointing task, we presented blindfolded and partially sighted participants with a set of virtual targets, one after another, and asked them to simply point a camera to where they thought the targets were.
The targets' pan and tilt angles w.r.t.\ the Tango's vertical plane, shown in \cref{fig:cam-coords}, were given to the participants through a spatial tone with varying pitch played via a set of bone-conducting headphones.
We use the latter to not interfere with the user's normal hearing function, which people with visual impairments typically rely on.
Because these headphones bypass the external structure of the ear responsible for localising a sound source's elevation, it is necessary to convey the target's tilt angle using another method.
Therefore, we opted for a tone with varying pitch to convey the angle adjustment required to point the camera in a particular direction.

@@

%####################################################
%REVIEW SECTION - MAKE SURE TO MENTION DIFFERENCE BETWEEN BLIND AND SIGHTED , sonification of 3d scenes in an electronic travel aid for the blind.
Spatial audio has also been considered to convey the direction of a target.
Experiments have previously determined that people are able to find the location of a stationary sound source with an error of $\pm$\SI{35}{\degree}, in both the pan and tilt dimensions~\cite{zwiers2001spatial}, for both early-blind and normally-sighted people.
However, in~\cite{lewald2013exceptional, lessard1998early} it was found that the blind have a clear advantage in localisation accuracy over sighted people when presented with more difficult tasks, such as targets in motion and narrow-band stimuli, with~\cite{lewald2013exceptional} reporting an average absolute localisation error of around \SI{10}{\degree} in the pan and tilt dimensions.
Other authors found that the minimum difference in the spatial sound's angle for the user to be able to perceive movement is approximately \SI{1.7}{\degree}. During the experiments, an audio speaker was physically manoeuvred to provide the participant with a spatialised sound tone~\cite{ashmead1998spatial}.
Researchers have also tried using simulated spatial audio to inform the user from which direction a sound is originating~\cite{holland2002audiogps, kammoun2012navigation, rebillat2009smart, menelas2010audio, wilson2007swan, zotkin2004rendering}.
In these works, a sound was played through a set of headphones and the source spatialised with a head-related transfer function (HRTF), tricking the listener into thinking the sound source was located at some arbitrary 3D location.
Various audition techniques and methods were then used to guide the user along a path.
Other authors have also provided a framework for evaluating the quality of this 3D sound in terms of its psychoacoustic properties~\cite{guastavino2004perceptual, nicol2014roadmap}.
There are also experimental results about how well users can find targets presented with spatial sound in the tilt and panning dimensions with normal headphones~\cite{katz2011spatial, zwiers2001spatial}, as well as work that has shown similar levels of localisation performance between normal and bone-conducting headphones in the pan dimension~\cite{macdonald2006spatial}.
However, to our knowledge no extensive work or experiments have been done to determine how well users respond to tilt adjustment instructions using a tone with \emph{varying pitch}, in particular when using bone-conducting headphones. 
%########################################################

@@

We use a set of bone-conducting headphones (see \cref{fig:tango-headphone}) that are placed externally on the user's head so that the system does not occlude the perception of real-world sounds and does not interfere with a user's normal hearing function.
In the future, our system will use multiple feedback modes to provide the user with navigation and obstacle avoidance instructions.
However, for this paper, we only considered the spatialised audio mode and its variation in pitch in order to determine its effectiveness in conveying a target's pan and tilt angles.

A diagram of the experimental system pipeline is shown in \cref{fig:pipeline}.
Here, the arrows indicate the direction of the flow of information.
When the user taps the Tango's screen, a new virtual target is generated and its coordinates are sent to the audio generation module, along with the Tango's current position and orientation.
The audio generator then produces a tone based on the difference between the Tango and the target's positions (in the Tango's global coordinate system) and sends it to the audio output channel, which plays it back to the user.
The WiFi recording module is constantly monitoring the different parameter values of the Tango and target's positions, as well as the system's output, and records it to a remotely stored datafile.

\begin{figure}
  \centering
  \includegraphics[clip=true, trim=50 120 80 50, width=0.4\textwidth]{figures/pipeline.pdf}
  \caption{A diagram of the individual system components and their communication pipelines. $F$ indicates a feedback signal and $P$ a pose signal. }
\label{fig:pipeline}
\end{figure}

@@

For the series of experiments performed in this work, only the audio feedback mode was used to interface with the user.
The audio component is responsible for conveying the 2D position of a target on the vertical plane w.r.t.\ the Tango's coordinate system, showed in \cref{fig:cam-coords}, in terms of pan and tilt angles.
Distance to the target can be rendered by adjusting the audio signal's amplitude as a function of the distance.
However, in this work, this was not implemented.
The audio signal is a sinusoidal sound wave that is continuously played (refreshed at $\sim$\SI{10}{\hertz}) to the user through bone-conducting headphones.
We select a sinusoidal wave because it is relatively simple to manipulate and analyse.
A richer and more natural tone will be implemented at a later stage of the project.

@@

The pan angle describes the angle which the user needs to rotate the camera vector around the vertical $z$-axis in \cref{fig:cam-coords}, i.e.\ how far the target is to the left or right of the user.
We use an HRTF to add a spatial element to the audio tone that the system plays to the user, making the tone sound like it is coming from the direction of the target.
Humans determine the pan direction by subconsciously analysing the time difference between a sound reaching both of your ears.
For example, if you hear a sound in your left ear first, you interpret the sound as coming from your left.
The greater the so-called inter-aural time difference~\cite{wightman1992dominant}, the greater the perceived angular distance to the sound source.

@@

The system adjusts the tone's pitch to convey the target's tilt angle w.r.t.\ the camera's current pointing vector, $C$, as shown in \cref{fig:cam-coords}. 
Here, a high pitch means the target is above the camera vector (i.e.\ the user should look up) and a low pitch means the target is below the camera vector (the user should look down).
This high/low association scheme is chosen because humans naturally associate high-pitched sounds with higher objects and lower-pitched noises with lower objects~\cite{pratt1930spatial, blauert1997spatial}.
We use a logarithmic, octave-based gain function for the pitch, since an increase in octave provides a distinct perceptible change while keeping the timbre roughly constant~\cite{shepard1964circularity}.

@@

The frequency is selected using Equation~\ref{eq:pitch-eq}, where $\theta$ is the tilt angle and $m$ and $c$ are parameters calculated using a linear equation (seen in Equation~\ref{eq:linear}) and the cut-off frequency values.
Here, $y_m$ is a cut-off frequency, $x_m$ its corresponding angle and $m$ the gradient for the setting, calculated as $m = \frac{\Delta y}{\Delta x}$.
The pitch gain functions are shown in \cref{fig:pitch-gain}.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.4\textwidth]{figures/pitch_gain_functions.png}
  \caption{The pitch gain functions for the 3 different gain settings. }
  \label{fig:pitch-gain}
\end{figure}

@@

For the experiments we used two groups of participants, with one including 42 sighted and blindfolded participants and the other 3 legally blind volunteers with varying degrees of vision impairment. 
They performed a series of experiments using our system with a pair of bone-conducting headphones.
The participants were recruited on a volunteer-basis and consisted of a diverse group of undergraduate students and professionals with ages ranging between 17 and 40 years (13 male, 32 female).
The participants also reported having no significant hearing issues or any other disability that could affect the experiment.
The participants participated in three experiments.
The first two experiments were performed to determine each participant's hearing characteristics and provide some context to the results generated during the target-search experiment.
These characterisation experiments will also indicate pre-existing biases our participant-base might have in the modes or dimensions we performed our third experiment on: the target search.

@@

The distance between the participant and the target is not considered here and the 3D targets are therefore generated at a constant distance of \SI{2}{\metre} from the participant.
Throughout the experiments, various parameters of the target and the participant are recorded and streamed in real-time to a laptop computer via a WiFi connection.
Prior to the target search experiment, the blindfolded participant group was given a few minutes without a blindfold to familiarise themselves with the system and confirm the target's location with their own eyes.
This was done verbally with the participants with limited vision.

@@

In this experiment, we evaluated a participant's ability to determine the lateral direction a sound is coming from.
To do this, we played a continuous \SI{512}{\hertz} sinusoidal tone to the participant through the headphones and applied an HRTF to spatialise and place its source to the participant's left or right.
The participant then had to select the direction the sound came from.
The longer the experiment lasts and the more correct guesses the participant makes, the closer the source moves to the centre-front of the participant, making it increasingly difficult to localise.

For this progressive increase in difficulty, a 2-up, 1-down step process is used~\cite{wetherill1965sequential,levitt1971transformed}, meaning that for every 2 correct answers, the distance to the centre halves, making the lateral direction harder to determine.
Conversely, the task becomes easier for each incorrect answer by doubling the sound source's distance from the centre.
We also use 2 different step sequences, one starting at a large distance (\SI{45}{\degree}) from the user and the other at the minimum distance (approximately \SI{1}{\degree}), giving an `easy' and a `hard' step respectively.
The terminating condition for the experiment is when the 2 step sequences converge to within 2 step ranges of one another for 3 consecutive guesses.
This gives a distance band within which the participant is capable of localising the sound source. Each participant performed this experiment three times.

@@

Here we determined a participant's ability to differentiate pitch, i.e.\ how well they can tell if a tone is high or low pitched.
We played 2 tones to the participants in succession with the second tone being higher or lower-pitched than the first.
The participants were then asked to select whether they interpreted the second tone as higher or lower-pitched than the first.

The first tone is randomly generated, while the second tone is generated by adding or subtracting some value from the first one.
The tone difference depends on how well the participant can tell the tones apart.
Like the spatial awareness experiment, a 2-up, 1-down step process is used: for every 2 consecutive correct answers, the pitch difference between the tones is halved, and the difference is doubled for every incorrect answer.
Two step sequences are again used here, one starting with a large pitch difference ($f_h=2^9=$ \SI{512}{\hertz}) between the tones and the other with a small difference ($f_l=2^1=$ \SI{2}{\hertz}).
The termination condition is when the two step sequences are within one octave of each other (i.e.\ $\log_2\frac{f_h}{f_l}=2^1$) for 3 consecutive answers.
Each participant performed this experiment twice.

@@

The final experiment is the main one and will answer the question we are most interested in: how well does a spatial tone with varying pitch direct a user to look in a specific direction and how do the parameters of this tone affect the user's performance in this task?

For this experiment, the participants were blindfolded where necessary and given a Tango device running an app purpose-written for this experiment.
When started, a set of virtual targets were presented one at a time to the participant on the Tango device.
Then, depending on the direction the participant was currently pointing the camera relative to the target's position, the Tango generated and played a tone via the bone-conducting headphones to indicate the pan and tilt angle adjustment required to point the camera to the target. %These instructions are spatialised tones with varying pitch: an HRTF indicates the pan direction (left or right) and the pitch indicates whether the participant should be looking up (high pitch) or down (low pitch) to find the target.

Once the participant pointed the camera toward the target, the system generated a neutral tone with a pitch of \SI{512}{\hertz}, which we used as the `on-target' pitch for all of our experiments.
The participants then had to manually confirm the direction they believe the target was after hearing the on-target tone and tap the screen.
At this point a new search target was presented and the search-process was repeated.
To minimise any speed/accuracy trade-off bias in the results, we asked the participants to search for the targets without placing particular emphasis on accuracy or time.

28 targets were presented to each participant per run.
The positions of these targets were randomly generated and equally spread across the 4 quadrants on the vertical plane to prevent a clustering of targets at the same location.
After every run of the experiment, the parameters controlling the tone's behaviour was adjusted.
In this case, the rate of change of the tone's pitch was adjusted to make the pitch increase at a lower or higher rate as a function of the tilt angle between the target and the participant's current pointing direction.
This was done to see whether, for example, a more rapid increase in pitch would help the participant find the target faster or more accurately.
The order in which the settings were set for each participant was randomised in order to minimise any learning effect the participant population might display between each experiment.

@@

We use two different metrics to compare the three different pitch gradient settings: the acquisition accuracy and search time.
The accuracy is given as the difference between the Tango's orientation at the time the participant confirmed they were on target, and the target's actual orientation.
We separate the results of the tilt and pan dimensions in order to see how the different pitch gradients affect a participant's pointing accuracy.

We also compare the performance of the three pitch gradient settings in terms of the time it takes each participant to find a target.
However, since each participant was presented with a different, randomly generated set of targets, a direct time comparison is not possible.
Instead, we use Fitts's Law~\cite{fitts1954information}, modified by MacKenzie for uncertain target sizes and noisy data~\cite{mackenzie1992fitts}, which states that there is a relation between the time it takes to find a target and its index of difficulty (the ratio between the distance to the target and its width).
It also provides a so-called `index of performance' that we can use as a metric to compare the results between the three configurations.

Here we briefly summarise the equations and quantities involved in our metric.
Fitts's Law is given by

@@

\noindent
where $t$ is the time it takes to find a target, $a$ and $b$ are constants determined through regression and $ID$ is a description of the difficulty of the target, given as logarithmic function of the ratio between the distance to the target and the target's width.
In our case, the targets have no width, since they are points in space, and we therefore use MacKenzie's modified form for $ID$, given by

@@

\noindent
where $\sigma$ is the standard deviation of the error data, taken as the angle between the participant's target selection and target's actual angular position.
Fitts's index of performance, $IP$, can then be calculated using

@@

\cref{fig:location-guesses} shows the results for the spatial awareness experiment in the pan direction where the participants had to determine the location of the sound they were played.

\begin{figure}
  \centering
  \includegraphics[clip, trim=100 20 150 100, width=0.4\textwidth]{figures/location_guesses.png}
  \caption{Histograms of the spatial awareness experiment in the pan dimension showing the proportion of correct/incorrect guesses per bin interval. }
  \label{fig:location-guesses}
\end{figure}

With the relatively low number of erroneous guesses, we see that the participants were accurate in guessing the location of the sound source.
This is further supported by the large number of samples at the minimum difference of \textasciitilde\SI{3}{\cm}, indicating that the participants reached this level more frequently and consistently.
We can therefore conclude that the participants had little difficulty determining the pan angle of a sound source.
These results are in line with what we expected and they are supported by literature which indicates that humans are very adept at localising a sound source, and this ability was apparent for the HRTF-generated pan locations.

The results recorded during the pitch discrimination experiment are shown in \cref{fig:tone-guesses}, where a bar plot is used to show the number of correct guesses for each tone difference level.
The latter is given in terms of the relative semitone difference between the tones which are obtained using the following relation:

@@

From \cref{fig:tone-guesses}, we can see that the guesses are normally spread around a 0 semitone difference.
As expected, a larger proportion of incorrect guesses occur around the smaller differences with the majority of samples for correct and incorrect guesses being recorded in the $[-0.25, 0.25]$ semitone difference level.
Assuming these differences are normally distributed, we fitted a cumulative distribution function (CDF) over each participant's set of results for their correct guesses and used the CDF's parameters to determine the semitone difference threshold, where the participant could no longer reliably tell two tones apart, for each participant.
See \cref{fig:cdf-semitone} for an example showing the actual CDF of a participant and its approximated normal CDF\@.
We set this threshold at 75\% of the correct guesses, since this represents a significant majority of samples.
We found that the median threshold is approximately 0.43 semitones for the sighted participants in a range of [$9.2\times10^-4$, 3.12] semitones.
We use the median given the skewness of the distribution.
The plot showing the participants' threshold distribution can be seen in \cref{fig:tone-threshold}.

%\begin{figure}
  %\centering
  %\begin{subfigure}[t]{0.5\textwidth}
\begin{figure}
  \centering
  \includegraphics[clip, trim=80 0 150 100, width=0.4\textwidth]{figures/tone_guesses.png}
  \caption{Results of the tone limit experiment. }
  \label{fig:tone-guesses}
\end{figure}

\begin{figure}
  %\end{subfigure}%
  %\begin{subfigure}[t]{0.5\textwidth}
  \centering
  \includegraphics[clip, trim=120 0 120 70, width=0.4\textwidth]{figures/cdf_tone_guesses.png}
  \caption{Cumulative distribution of the number of correct guesses for one participant. }
  \label{fig:cdf-semitone}
\end{figure}
  %\end{subfigure}
  %\caption{Histograms of the tone limit experiment showing the proportion of correct/incorrect guesses per bin interval. }
%\end{figure}
